UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-06-05 ms/main.tex
couple of small edits

Here we will focus on nonparametric approaches to visualize relatedness patterns among individuals within populations. If we consider unphased single nucleotide polymorphism (SNP) data, an individual genome can be represented as a sequence of integers corresponding to the number of alleles carried by the individual at each of the $L$ SNPs for which genotypes are available, with $L$ typically larger than $100,000$. Since each individual is represented as an $L$-dimensional vector, dimension reduction methods are needed to visualize the data.

@@

We found that pre-processing the data with PCA allowed for considerable time savings. Groupings on ethnicity formed slowly until reaching a stable number around 10 to 15 components. Geographical patterns in the UKBB continued to appear even up to 40 components, as visible in figures~\ref{fig:supp_montage_ukbb_ns} and \ref{fig:supp_montage_ukbb_ew}. In our experience, we find it useful to include as many PCs as are available and computationally feasible.

>>>>

Here we will focus on nonparametric approaches to visualize relatedness patterns among individuals within populations. If we consider unphased single nucleotide polymorphism (SNP) data, an individual genome can be represented as a sequence of integers corresponding to the number of alleles carried by the individual at each of the $L$ SNPs for which genotypes are available, with $L$ ranging from hundreds of thousands to hundreds of millions. Since each individual is represented as an $L$-dimensional vector, dimension reduction methods are needed to visualize the data.

@@

We found that pre-processing the data with PCA allowed for considerable time savings. Groupings on ethnicity formed slowly as PCs were added until reaching a stable number around 10 to 15 PCs. Geographical patterns in the UKBB continued to appear even up to 40 components, as visible in figures~\ref{fig:supp_montage_ukbb_ns} and \ref{fig:supp_montage_ukbb_ew}. In our experience, we find it useful to include as many PCs as are available and computationally feasible.
